P4: Top 10k Best Selling Video Games - Exploration in R by Whitney King
===============================================================================

> **Tip**: One of the requirements of this project is that your code follows
good formatting techniques, including limiting your lines to 80 characters or
less. If you're using RStudio, go into Preferences \> Code \> Display to set up
a margin line to help you keep track of this guideline!

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

#install.packages('ggplot2')
library(ggplot2)

#install.packages('ggthemes', dependencies = TRUE)
library(ggthemes)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
vgdata <- read.csv('vgsales.csv')

#subset(vgdata, Genre == 'Puzzle')
```

# VGCharts: Top 10,000 Best Selling Video Games Globally
## Data Exploration in R
#### By: Whitney King

Data obtained on 5/22/2017


## Introduction

This data exploration will take an in depth look at data scraped from the 
VGChartz.com charts for regional and global video games sales (by millions 
USD). This data was obtained on 5/22/2017 using a Python3 script, and 
importing BeautifulSoup to parse out the HTML data.

After the dataset was scraped from the table on 
[VGChartz website](/http://www.vgchartz.com/gamedb/), it was then limited 
to the top 10,000 rows, and formatted using a dataframe before being output to
CSV for use in R.

The dataset is structured as follows:
* **Rank** *(int)*;
  Primary Key/Unique Identifier for each game in the list,
  ranked best-selling to worst-selling
* **Name** *(str)*;
  Title of the video game
* **Platform** *(str)*;
  Console/Platform game was released on
* **Year** *(int)*;
  Year game was released
* **Genre** *(str)*;
  Genre/Category of the game title
* **Publisher** *(str)*;
  Publisher of the video game
* **NA_Sales** *(double)*;
  Sales in millions (USD) in North America
* **EU_Sales** *(double)*;
  Sales in millions (USD) in Europe
* **JP_Sales** *(double)*;
  Sales in millions (USD) in Europe
* **Other_Sales** *(double)*;
  Sales in millions (USD) in other regions of the globe
* **Global_Sales** *(double)*;
  Total sales (USD) globally

 
## Univariate Plots Section

It will be important to understand a little bit about the data in this dataset
prior to working with it.

```{r echo=FALSE, VGData_Columns}
names(vgdata)
```
The ```names()``` function is a good starting point, since it generates a list
of column names. This is a simple way to ensure the data ingested in the 
expected format.


```{r echo=FALSE, VGData_Overview}
str(vgdata)
```
Converting the dataframe to a string shows a basic breakdown of how the data 
was ingested. The Year column is showing as a factor due to the 'N/A' values. 
It should be a continuous variable, so this will need to be taken into 
consideration as the analysis progresses.


```{r echo=FALSE, VGData_Summary}
summary(vgdata)
```



```{r echo=FALSE, Univariate_Plots, fig.height = 100, fig.width = 200}
dYear <- as.numeric(as.character(vgdata$Year)) #convert string to continuous variable

ggplot(vgdata, aes(dYear)) +
geom_histogram(binwidth = 2,
               aes(fill = ..count..),
               col = 'darkgreen',
               alpha = .6) +
theme_set(theme_minimal(10)) +
theme(axis.text.x = element_text(angle = 90, hjust = 1),
      aspect.ratio = 1) +
labs(title = 'Histogram: Video Games Released by Year',
     subtitle = 'Worldwide Sales',
     x = 'Year',
     y = '# Video Games Released') +
scale_fill_gradient("Count", low = "red", high = "green") +
scale_x_continuous(breaks = seq(1980, 2020, by = 2))
```


> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!